---
title: "assignment_1"
output: html_document
---

```{r read_csv, echo=TRUE}
rawdata <- read.csv("/Users/heathermurray/Downloads/activity.csv", header = TRUE)
require(dplyr)
library(dplyr)
```
#### Days vs. Steps
```{r, make_hist1}
remove_na <- filter(rawdata, steps !="NA")
groups <- group_by(remove_na, date)
summary <- summarize(groups, mean_steps = mean(steps, na.rm = TRUE), median_steps = median(steps, na.rm = TRUE))
as_frame <- data.frame(summary)
mean_steps <- select(as_frame, date, mean_steps)
median_steps <- select(as_frame, date, median_steps)
as_date <- as.Date(as_frame$date)
plot(as_date, mean_steps$mean_steps, type = "l", xlab = "Date", ylab = "Steps")
```
#### The mean and median 
```{r mean_median}
print(mean_steps)
print(median_steps)
```

```{r make_plot1, fig.height=20, fig.width=20}
group_interval <- group_by(remove_na, interval)
mean_interval <- summarize(group_interval, mean_steps = mean(steps, na.rm=TRUE))
intervals <- mean_interval$interval
steps <- mean_interval$mean_steps
plot(intervals, steps, type="l", xaxt="n")
axis(1, at=c(seq(0, 2355, by=100)), labels=TRUE)

```

#### Counting the number of missing values
```{r number_na}
num_na <- nrow(rawdata)-nrow(remove_na)
print(num_na)
```

#### Replacing the NA values with the mean of the interval gives you this
```{r replace_means}
replace_means <- rawdata %>%
  group_by(interval) %>% 
  mutate(steps = replace(steps, is.na(steps), mean(steps, na.rm=TRUE)))
plot(replace_means$interval, replace_means$steps, type = "l", xlab="interval", ylab = "mean steps")
axis(1, at=c(seq(0, 2355, by=100)), labels=TRUE)
```

#### Dividing the week
```{r days_of_the_week}
dates <- as.Date(replace_means$date)
days <- weekdays(dates)
replace_means["weekdays"] <- days
mon_fri <- subset(replace_means, weekdays !="Saturday"| weekdays != "Sunday")
weekends <- subset(replace_means, weekdays == "Saturday" | weekdays == "Sunday")
mean_weekday <- summarize(mon_fri, steps_mon_fri = mean(steps, na.rm=TRUE))
mon_fri_mean <- mean_weekday$steps_mon_fri
mean_weekend <- summarize(weekends, steps_weekends = mean(steps, na.rm=TRUE))
weekend_mean <- mean_weekend$steps_weekends
```

```{r plot_days}
par(mfrow=c(1,2))
plot(mean_weekday$interval, mon_fri_mean, type = "l", xlab = "interval", ylab = "average weekday steps")
plot(mean_weekend$interval, weekend_mean, type = "l", xlab = "interval", ylab = "average weekend steps")
```
There are more periods of activity on weekends than weekdays. On weekdays the greatest number of steps is concentrated around the 800-1000 second interval. On the weekend there are multiple peaks between the 800-2000 intervals. 